A thesis draft

### 2018-11-29 res/sections/10-ProjectAnalysis.tex
Add DNSMasq software description (#51)

@@

@@

@@

>>>>

\section{Architecture overview}

During the test-bed development the software architecture diverged from the
original one, experiencing different evolutions and modifications.

\subsection{VIBES architectural description}

In the original VIBES architectural design, based on leveraging virtualization
technologies, is possible to identify different macro-components, each of these
specialized in fulfilling a specific goal. For the ETSI standards, the Network
Function Virtualization (NFV) design should be composed as follows:
\begin{figure}[t]
 \centering \includegraphics[scale=2]{etsi_arch}
 \caption{A schematic image representing the high level NFV architecture
   proposed by ESA.}
 \label{chap:prjan:img:etsi_arch}
\end{figure}

\begin{itemize}
 \item \textbf{Network Function Virtualization Infrastructure (NFVI)}: it's the
   heart of the whole computation component, and here three sub-domains can be
   identified: \todo{Expand this section}
\begin{itemize} 
 \item \textbf{Hyper-visor domain} that is composed of the Virtual Compute, the
   Virtual Storage, and the Virtualization Layer. These components are used to
   virtualize the computational and storage resources, providing a layer for
   accessing it.
 \item \textbf{Compute domain} composed by the Physical Compute and the Physical
   Storage. These are the real resources which are virtualized via software.
 \item \textbf{Infrastructure domain} which has the Physical Network, and its
   virtualized counterpart as made available by the Virtualization Layer.
\end{itemize}

In conventional virtualization systems these resources are usually separated
from the host operating system, providing better isolation but worst
performance. The ETSI infrastructure, instead, tries with container
virtualization technology to push for a better NFV responsiveness, removing the
additional operating system (called ``guest'' OS) required by the usual
hyper-visor-oriented approach.

To realize the required components, specific tools were suggested in the VNF
technical proposal: regarding the NFV (network infrastructure) domain
Docker-Compose was proposed as a viable solution, while to store PEP instances
docker swarm was described as a possible candidate.

We performed an accurate analysis of the tools chosen based on their maturity,
on the community and on the technical support offered (e.g. user manuals,
developer documentation), which, at the end, committed us to choose different
tools from the suggested ones. A detailed description of the architectural
implementation can be found in chapter~\ref{chap:archimpl}.\todo{Include prof.
  survey about community components here}

\item \textbf{VNFs} this are the core components that perform packets
  manipulation. This components are design to be lightweight services able to
  elaborate a great amount of packets per seconds. Manipulation can be of two
  typologies: stateful or stateless. In the first case, a state is maintained,
  and the VNF can perform heuristics about data coming through it (especially
  useful to detect patterns in the data flow, e.g. DDos attacks). In the second
  case, no state is maintained, and every packet is treated without any
  particular assumption. Following the RFC 7665, the VNF are also called Service
  Function (SF)\footnote{From now on the two terminologies will be use in a
    interchangeable way~\cite{medhat2017service}.}. Additional information can
  be found in Section~\ref{chap:prjan:sfc}.

\item \textbf{NFV Management and Orchestration (MANO)} also known as NFV
  Orchestrator (NFVO) is responsible for the end-to-end management and
  orchestration of network services (NS), also called SFC. The MANO has
  different responsibilities, such as managing the
  creation/destruction/modification of the SFCs, that achieves interacting with
  the VIM~\cite{nguyen2017sdn}. Finally, the MANO can be integrated with
  Operational and Business Support Systems (OSS/BSS).
\end{itemize}

\section{SFC}
\label{chap:prjan:sfc}

\todo{Here it's missing an introduction about SFC and its terminology, need to 
include parts of RFC 7665}

Even though SFCs and VNFs are part of the ETSI architecture, the complexity 
behind was so high to require particular care and a separated study. Indeed, 
the  SFCs are the core of the whole project: packet transmission and 
elaboration is made here. Challenges regarding this topic are building a 
system that is flexible enough to allow dynamic packet routing, support 
scalability and provide methods to recovery from failures. With this challenges 
packet transmission revealed to be a fundamental point to choose carefully.

\subsection{Packet transmission strategies}

When a packets are sent from a sender $S$, to a receiver $R$, they start
travelling to the backbone receiving appropriate in-hardware packet processing
before being received by $R$. During this elaboration, it is important to
maintain the transparency of the connection, so that the network users are not
aware of the packets elaboration at all. To perform that, many solutions are
available: TUN/TAP packet encapsulation, IP Encapsulation within IP or TCP
session recreation (know also as TCP split).

TUN/TAP and IP Encapsulation within IP (initially defined in RFC 2003 with the
aim to deliver IP packets over Mobile IP for mobile networks) maintain the
connection transparency encapsulating the packet in another one.
\begin{figure}[t]
  \centering \includegraphics[scale=0.5]{IPoverIP}
  \caption[IP Encapsulation within IP packet schema]{A schema of IP
    Encapsulation within IP. Is noticeable how the original packet becomes the
    payload of a new one, thus being encapsulated. All the information
    regarding the original packet remains untouched: when this kind of packet
    gets send from a machine to another, the OS will remove the new packet
    header and will present to the user-space program receiving the data its
    payload: a IP packet. Packets encapsulation can be applied with higher
    layers too, thus allowing for TCP and UDP packet encapsulation.}
  \label{chap:prjan:img:ip_over_ip}
\end{figure}

In this way, the original packet remains preserved and it becomes the new packet
payload, that can be elaborated and modified accordingly, until it gets
decapsulated and sent to $R$ eventually. When $R$ receives the packet, it is not
be aware of the packet elaboration, since all the original headers were
preserved (or slightly modified). In particular, IP Encapsulation within IP
allow the packets to make intermediate destinations that otherwise would not be
selected. TCP session recreation, instead, ``split'' the TCP session in two
endpoints: an Ingress, that manage the connection with the client and an Egress,
that communicate with the packet receiver creating a multi-overlay-hop path
where for each hop there is an possible independent TCP connection. In order to
achieve that, a session table needs to be established and continuously updated,
while TCP headers need to be accordingly modified to allow data transmission
through the Ingress and Egress points. Nonetheless, care should be given to
packets path: in order to avoid ill-calculated congestion windows, the incoming
packets need to follow the same (or an equivalent one) path for the same
instantiated connection. This phenomenon is based on the intermediate proxies
that create the route: it is the case, for example, when an intermediate node
sends back a spoofed ACK before the original packet is delivered to the real
consignee. In this scenario, the sender, unaware of the proxy, will miscalculate
the sending window based on the too low RTT. Other problems of TCP splitting are
reliability and security, especially because the end to end TCP logic is no
longer maintained: a server failure may cause to an unaware client to believe
that all the packets have reached the destination. TCP session recreation seems
burdensome and it doesn't seem to add any significant benefit at the first
sight, but it offers the flexibility to perform extensive packet manipulation:
since the packet its recreated every time at the Ingress and Egress points,
while in the Network Function Virtualization Infrastructure (NFVI), the original
packet can be stripped of its headers (that could be saved in the session table)
to slightly improve internal packet transmission. \todo{Insert TCP session
  recreation schema} On top of that, it offers the possibility to use different
TCP flavours, hence increasing the communication speed in heterogeneous networks
(e.g. in case of satellite connections TCP Hybla can be employed, while it is
known that HighSpeed TCP gives best results when applied in optical fiber-based
connections).

During the test-bed development we first tried the TUN/TAP solution, that
revealed to be not completely transparent and with major performance drawbacks
(i.e. packet transmission in a LAN connection reached latencies of 50ms)
\todo{Explain the constraint of point-to-point connections, while we need
  dynamic SFC)}, so we choose to opt for a TCP session recreation solution,
implementing a little back-end with the aim of saving in a persistent storage
the SFC the packet has to perform and additional metadata.

\section{Available technologies in the market}
\label{chap:prjan:sec:tech}

Before starting any effective work on the project it has been decided to 
perform an analysis of the possible technologies available in the market, 
performing a choice between them and exploiting their possible ``weak 
points''. Proceeding with a top-down approach, the frameworks were chosen from 
the one that had the biggest role in the test-bed to conclude with the 
less-important components.

\subsection{NFVI}

A critical component, as the MANO one, is the NFVI. Without it, network 
resources can't be allocated, and since our project was focused on dynamically 
allocating containerized applications on cloud platforms, we decided to start 
analyzing the right framework for this component first.

@@

\subsubsection{Harbor}
\label{chap:prjan:sec:tech:sub:MANO:sub:harbor}
Started as a VNFM (Virtual Network Function Virtualization Manager) to offer a
bridge for Openbaton to interact with Kubernetes, during the project it has been
expanded to become a complete replacement of the original product, adding the
possibility to manage Network Services and to update the routing tables
contained in the Roulette Controller (the tool description can be found below
in~\ref{chap:prjan:sec:tech:sub:SFC:sub:roulette}). Harbor doesn't have a web
interface, but it has a RESTful API that allows users to interact with it using
CRUD operations. It offers functionalities similar to Openbaton: possibility to
save and retrieve VNF and NS definitions (basically, Kubernetes YAML
deployments). When an NS is saved, a consistency check is performed to assure
that all the VNFs have already been defined in Harbor. On an NS deployment, with
a mechanism similar to the Openbaton one, all the associated VNFs are set up
too, making the whole operation of launching a NS atomic from the user
point-of-view. What Openbaton doesn't do is a smart resource utilization:
Harbor, in fact, during a NS deletion, it checks if any related VNF remains
orphan (which means without any NS using it). In that case the VNF will enter in
a state of pruning: after five minutes, if the VNF has not be re-utilized by any
other NS it will be deleted. This mechanism assure faster deployments in
environments where NSs are continuously taken up and down. \todo{I've made an
  Harbor implementation schema, I dunno if its useful to put it here}

@@

\subsubsection{Roulette Controller}
\label{chap:prjan:sec:tech:sub:SFC:sub:roulette}
Dynamic communications between VNFs would not be possible without a third
component acting as a route controller. Roulette fulfills this role, providing a
back-end where the single VNF can ask the next packet hop. Indeed, routes must 
be
submitted by another component, that we identified as Harbor (component
description available above in
sub-section~\ref{chap:prjan:sec:tech:sub:MANO:sub:harbor}). After a NS
deployment, Roulette gets instructed with a new route describing a NS. Every NS
has a proper ID (called \verb!sfid!), and thus when the classifier checks a
packet typology, it labels that packet with the specific NS ID (for a detailed
description about the SFC header see Section~\todo{Add section/chapter/paragraph
  ref!}). Finally, thanks to this label, every VNF knows the packet's chain and
it will correctly forward the it to the next VNF.

\subsubsection{Astaire}
As is possible to denote, a VNF doesn't only process the incoming packets, 
but it has to manage the communications and the related forward operations. To 
possibly ease part of the transmission problem and to make it easier to create 
a 
VNF without requiring the whole knowledge about SFC, it has been decided to 
build a framework, Astaire, that could keep the SFC related details decoupled 
from the core of the VNF, in other words, the entity that actually elaborates 
the packet. Built using an event-based pattern, it presents hooks that allow 
other software written in different languages from the Astaire one (e.g. Java), 
to be called dynamically at runtime in order to process the packet. The result 
of this elaboration gets returned to Astaire, that performs the beforehand 
cited operations of VNFs communications and forwarding.

\subsubsection{Ironhide}
As described in chapter~\ref{chap:archimpl}, the whole framework need to have 
and entry and an exit point. These, defined as \textit{Ingress} and 
\textit{Egress} points, are performed by Ironhide. Ironhide acts like a proxy, 
hiding the whole packet processing to the sender $S$ and the receiver $R$. When 
a client $S$ performs a request to $R$, the packet actually goes through 
Ironhide, who has an internal classifier that it is able to recognize the type 
of connection and to choose the right NS the packet has to go through. This 
process in fundamental since it keeps the connection between $S$ and $R$ up, 
for both TCP and UDP based communications.

@@

\subsubsection{Dnsmasq}
\label{chap:prjan:sec:tech:sub:other:sub:dnsmasq}
Dnsmasq is an open source project, started in 2001~\cite{dnsmasqweb}. Its
features includes the possibility to serve as DNS, DHCP or router advertisement
and network boot. Dnsmasq is designed to have a small footprint memory, and it
is suited for medium to small network infrastructures. Along with the DNS
server, it has the possibility to perform request caching, to speed up
additional requests of the same type. On top of that, it provides different
compilation options in order to eliminate unused features at compile time.
